---
title: People Are Naming Their Children After Creed Bratton
author: Daniel Baggett
date: '2019-06-09'
slug: kramer
categories: []
tags: []
authors: []
---

```{r library chunk, echo=FALSE, include=FALSE}
library(tidyverse)
library(babynames)
library(knitr)
library(stringr)
library(janitor)
library(blogdown)
library(rmd)
library(tinytex)
```

If you've used the internet this year, you're likely aware of the fact that people are naming their children after popular characters from HBO's Game of Thrones. 

Maybe you've seen a chart like this:

```{r Daenerys All, echo = FALSE}
Daenerys_all <- babynames %>% filter(name == "Daenerys" & sex == "F")
ggplot(Daenerys_all, aes(x = year, y = n, color = name)) + 
  geom_line(size = 1.0) +
  labs(title = "Sweetie, you are named after the Mother of Dragons") +
  ylab("Number of Children") +
  xlab("Year")
```

In 2016, more than 100 children were named after the Game of Thrones character Daenerys. 

**No one** named their child Daenerys before show's debut in 2011. It's safe to conclude that the show was responsible for the child's name and identity going forward in life. 

Working from the ever-popular [babynames dataset](https://www.kaggle.com/kaggle/us-baby-names), I wanted to see if people were naming their children after characters that closely follow the Game of Thrones pattern.

Did anyone name their child [Seven like George wanted to in Seinfeld?](https://www.youtube.com/watch?v=NRUdaWZ4FN0) What about [Beans from Even Stevens,](https://www.imdb.com/name/nm0493037/) or [Phoebe from Friends?](https://en.wikipedia.org/wiki/Phoebe_Buffay)

#### Criteria:
  
A. Less than 50 children per year shared a character's name before the respective show's debut.  

***
### The results are divided into three groups. 

1. Honorable Mentions - Barely missed 25 per year mark. 
2. Truly Odd  **< 50** uses per year before the show's debut.
3. Too weird for any parent.

***
### Honorable Mention Category
```{r honorable mentions, echo=FALSE}
honorable_mentions <- babynames %>% filter(name == "Phoebe" & sex == "F" | name == "Sheldon" & sex == "M" | name == "Rory" & sex == "F" | name == "Lorelei" & sex == "F", year >= 1950)

honorable_mentions_tot <- honorable_mentions %>% group_by(name) %>% 
  summarise(total = sum(n))

ggplot(honorable_mentions, aes(x = year, y = n, color = name)) + 
  geom_line(size = .9) +
  labs(title = "Honorable Mentions") +
  xlab("Year") +
  ylab("Children Named")
```

1. Lorelei, Gilmore Girls (2000 to 2007)
2. Phoebe, Friends (1994 to 2004)
3. Rory, Gilmore Girls (2000 to 2007)
4. Sheldon, Big Bang Theory (2007 to It Will Never End)

The popularity of the Big Bang Theory spawned a prequel of a show inside of a show that is **still running.** I figured Sheldon would be a lot higher. 

Even though Phoebe missed the cut, the data shows that Friends has transformed Phoebe into a moderately popular named when compared to it's past frequency. 

***

### Truly Odd. 
```{r Odd Names, echo=FALSE}
Odd_Names <- babynames %>% filter(year >= 1970, name == "Creed" & sex == "M" | name == "Gunther" & sex == "M" | name == "Kramer" & sex == "M" | name == "Seven" & sex == "M")
 
ggplot(Odd_Names, aes(x = year, y = n, color = name)) + 
                  geom_line(size = 1.0) +
                  labs(title = "Truly Odd Names from TV Shows") +
                  xlab("Year") +
                  ylab("Number of Children Named")
```

1. Creed, The Office (2005)
2. Gunther, Friends (1994 - 2004)
3. Kramer, Seinfeld (1989 - 1998)
4. Seven, Seinfeld (1989 - 1998)



Yes... That **Creed Bratton** line is accurate!
<br>
Yes... People were, and still are, naming their children "Seven!" 

To further prove the point, here is the average number of names, per year, adjusted for the show's time on television. 

```{r Per Year, echo = FALSE}
meanCreed <- babynames %>% filter(name == "Creed" & sex == "M", year >= 2005) %>% 
  group_by(name) %>% 
  summarize(mean = mean(n)) 

meanSeven <- babynames %>% filter(name == "Seven" & sex == "M", year >= 1989) %>% 
  group_by(name) %>% 
  summarize(mean = mean(n)) 

meanKramer <- babynames %>% filter(name == "Kramer" & sex == "M", year >= 1989 & year <= 1996) %>% 
  group_by(name) %>% 
  summarize(mean = mean(n)) 

meanGunther <- babynames %>% filter(name == "Gunther" & sex == "M", year >= 2005) %>% 
  group_by(name) %>% 
  summarize(mean = mean(n)) 

truly_odd_means <- rbind(meanGunther, meanKramer, meanSeven, meanCreed)

truly_odd_means %>% mutate(name = fct_reorder(name, mean)) %>% 
  ggplot(aes(name, mean, fill = name)) + 
  geom_col(width = .8) +
  coord_flip() +
  labs(title = "Average per year after show debut.") +
  ylab("Average Per Year During Show Run") +
  xlab("Names")

```


I then figured that Creed must have been a very popular name sometime in the past and now it is just coming back into style.
```{r Bonus Creed, echo=FALSE}
Creed_All <- babynames %>% filter(name == "Creed" & sex == "M")
ggplot(Creed_All, aes(x = year, y = n)) + 
  geom_line() + 
  labs(title = "Children Named Creed Since 1880")
```
To be fair, Creed was on the rise before The Office debuted in 2005; however, It's safe to say that some of the inspiration is from the character depicted on the show.  

### Too weird for literally anyone. 
No one, and I mean no one, has named their child after [Beans](https://www.youtube.com/watch?v=NXOISYS5ubY&t=29s) from Even Stevens. When Shia LaBeouf filmed himself watching himself in all of the movies he has ever starred in, he cried [in only one](https://www.youtube.com/watch?v=NXOISYS5ubY&t=29s).


