---
title: Looking At The Largest 100 Churches In America
author: Daniel Baggett
date: '2019-07-26'
slug: exploring-the-largest-100-churches-in-america
categories:
  - church
tags: []
authors: []
---
```{r include=FALSE}
library(tidyverse)
```

```{r include=FALSE}
top_coded <- read_csv("~/R/top_100_churches/top_coded.csv")
```

Each year, [Outreach Magazine](https://outreachmagazine.com/) and [Lifeway Research](https://lifewayresearch.com/) colloborate to produce the "Outreach 100." The two companies survey churches, verify data, and then compile lists of the fastest growing churches and the largest churches in America. They've recently made some of this information available for free on their [website.](https://outreachmagazine.com/outreach-100-list.html#/list/2015/all)

I wanted to visualize some of their data. Every church on the "Outreach 100" has an HTML Table associated with it. Using R I was able to arrange the HTML tables into something more useful. 

### For Example:

```{r echo=FALSE}
top_coded %>% 
  filter(year == 2018) %>% 
  top_n(25, attendance) %>% 
  mutate(church = fct_reorder(church, attendance)) %>% 
  ggplot(aes(church, attendance, fill = church)) +
  geom_col() +
  coord_flip() +
  labs(title = "25 Largest Churches in America",
       subtitle = "2018",
       caption = "Outreach Magazine", 
       x = "Church", 
       y = "Attendance") +
  theme(legend.position="none") 
```

A quick glance at the list above sparks a lot of questions that I'd love to answer here; however, *I will only work with data that is available for free* on their [website](https://outreachmagazine.com/outreach-100.html). I'll work though a few exploratory visualizations I made. If you're interested in the code, you can find it on github. 

<br/>

## 5 Questions Answered
#### 1. How do the 100 largest churches break down in regards to affiliation?

```{r echo=FALSE}
top_coded %>% 
  filter(year == 2018) %>% 
  top_n(100, attendance) %>% 
  count(affiliation) %>% 
  mutate(affiliation = fct_reorder(affiliation, n)) %>% 
  ggplot(aes(affiliation, n, fill = affiliation)) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  labs(title = "100 Largest Churches By Affiliation",
       subtitle = "2018", 
       x = "Affiliation", 
       y = "Number of Churches in Top 100 Attendance List",
       caption = "Outreach Magazine") +
  theme(legend.position="none") 
```

* I anticipated nondenominational churches to dominate the 2018 list; still, I didn't expect it to be this overwhelming. Also, I visited the websites of 10-15 churches that **do** declare their affiliation. I was looking to see how, or if, they communicated their respective denomination. Many of the churches on the list make their respecitve affiliation difficult to find. 

<br/>

### 2. Are large churches still growing?

```{r echo=FALSE}
raw_growth_four_years <- top_coded %>% 
  group_by(church, affiliation) %>% 
  summarise(raw_growth = sum(growth_by_number)) %>% 
  ungroup() %>% 
  arrange(desc(raw_growth)) %>% 
  filter(raw_growth > 3000)

#Even though I filtered out the levels before - this was the only way to get rid of the extra ones. 
raw_growth_four_years$church <- fct_drop(raw_growth_four_years$church)


raw_growth_four_years %>% 
  mutate(church = fct_reorder(church, raw_growth)) %>% 
  ggplot(aes(church, raw_growth, fill = affiliation)) +
  geom_col() +
  coord_flip() +
  labs(title = "Raw Numerical Growth",
       subtitle = "2015 - 2018", 
       y = "People Added", 
       x = "Church") +
  theme(legend.title=element_blank())
```

* [Church of the Highlands](https://www.churchofthehighlands.com/) stands out in nearly every graphic I make. They have added 15,000 people in four years. It's kind of staggering.  


* Alabama's population is approximately [4.8 million.](https://factfinder.census.gov/faces/nav/jsf/pages/community_facts.xhtml?src=bkmk) This means that Church of the Highlands represents *nearly 1%* of Alabama's population and they are still growing rapidly. No other church on the list comes close to this kind of population / state ratio.  


* The data above only includes the last four years because it is all that is available via Outreach. I also filtered the data down to churches that added at least 3,000 people in the last four years to make it managable to read. 



<br/>

### 3. Two looks at multisite models. 

```{r echo=FALSE}
top_coded %>% 
  top_n(100, attendance) %>% 
  filter(year == 2018 & locations >= 3) %>% 
  mutate(church = fct_reorder(church, locations)) %>% 
  ggplot(aes(church, locations, fill = church)) +
  geom_col() +
  labs(title = "Largest Multisite Churches In America", 
       subtitle = "Number of Campuses Per Church - 2018",
       x = "Church",
       y = "Number of Locations") +
  coord_flip() +
  theme(legend.position="none")
```

```{r echo=FALSE}
top_coded %>% 
  filter(year == 2018 & attendance > 12500) %>% 
  mutate(location_ratio = attendance / locations) %>% 
  mutate(church = fct_reorder(church, location_ratio) %>% fct_rev()) %>% 
  ggplot(aes(church, location_ratio, fill = church)) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  guides(fill=guide_legend(title=NULL)) +
  labs(title = "Attendees to Campus Ratio", 
       subtitle = "Churches Over 12k - Year 2018", 
       x = "Churches Over 10k In Attendance", 
       y = "Attendees Divided By Number Of Campuses") +
    theme(legend.position="none")
```

* Obviously, the second graphic does not represent the exact amount of people at **each** campus. I wanted to get an idea of each church's philosophy in regards to when they thought it was time to plant a new church, or when they believed it was time to add a satellite campus. 

* I also wanted to see the rate at which churches were adding campuses. The dataset does not allow me to see number of campuses added year by year. 

#### Final Thoughts

* A bevy of nondenominational churches are quickly creating networks of smaller churches that share core beliefs, values etc. so that people can experience something similar whereever they go. A cynic would say that nondenominational churches are slowly creating... denominations. I think the situation requires more nuance. 

<br/> 
















